\Large \bf Cyclone: Replication Middleware for NVM Applications

### 2016-09-26 paper/paper.tex
More tweaks to paper.

\subsection*{Abstract}
The introduction of non-volatile directly attached memory in datacenters - in
the form of NVDIMMs and new memory technologies allows efficient data structures
to be built on non-volatile heaps that provide good performance coupled with
durability. However high availability remains a problem for these primarily
single machine solutions. While transactional interfaces built over RDMA capable
networking hardware offer a viable alternative, such hardware is often
unavailable in datacenters. Cyclone offers a solution to this problem by
providing transparent replication middleware that can be used with Intel's Non
Volatile Memory Library (NVML) to construct durable data-structures that map
keys to values.  Cyclone is designed around an optimized userspace networking
stack to reduce latency and can scale its consensus protocol across cores and
network ports to provide good throughput under load. We show that Cyclone can
replicate as many as 8 million updates a second using just 4 commodity 10 GigE
ports on a machine.

@@

\section{Single Node Programming Model}
Cyclone assumes a concurrent durable application on a non-volatile heap as the
starting point. A number of libraries and filesystems make this possible
today~\cite{nvml, dax, pmfs, mnemosyne, nvheaps, cdds}. Cyclone currently
supports applications built on top of Intel's Non Volatile Memory Library
(NVML~\cite{nvml}). We describe NVML's programming model in this section to
provide adequate background for the rest of this paper. NVML requires 
that code that makes changes to non-volatile heap is enclosed in a crash
consistent transaction such that either all of the transaction executes or none
of it executes in the event of power failures.
%This is critical to maintaining consistency as the data structure, being
%durable, must survive across power failures.
Figure~\ref{fig:example} illustrates the use of NVML to build a persistent
linked list. Updates to the linked list are enclosed in a NVML transaction to
avoid corruption due to power failure. It is also worth noting that NVM
libraries such as NVML usually allow remapping non-volatile heaps to a different
virtual memory address range across restarts (e.g., using {\tt TOID} macros to
implement fat pointers in NVML).

@@

Asynchronous replication provides good performance and we believe it would be
the common case for RPC calls replicated via Cyclone. However the program might 
on occasion require successful replication to be a precondition to exection.
A case where we find this necessary is
\textit{output commit}~\cite{output_commit} - where the RPC call has
external side effects (e.g., communication with other machines in the cluster)
and therefore crash consistency is insufficient to roll back the execution of
the RPC call in the event that replication fails. We therefore provide users the
ability to declare at the client side that an RPC call should execute
synchronously at the server side. Cyclone would then ensure that the RPC
call is successfully replicated before it attempts execution on any of the
replicas.

@@

We exploit the fact that DPDK allows prepending data to a packet
without physically moving the data in a packet buffer. The received packet
contains only the RPC call from the client. We prepend a RAFT related header
(containing the previous log index, previous log term, leader's term and
leader's commit point~\cite{raft}) to the packet buffer itself. To this RAFT header 
we then prepend an IPv4 header. We assign pre-determined values to the
IP address fields in order to exploit 5-tuple steering at the receiving NIC for 
the placement of the incoming packet in the correct queue.
This composite packet buffer is then sent \emph{unchanged} to every replica.
To properly route the packet through switches, we allocate a small buffer
containing an ethernet header for each follower replica and \emph{chain} the
buffer in front of the prepared packet - thereby ensuring that we do not need
to change the contents of the packet itself for every replica.

@@

Cyclone provides at most once semantics. It supports a large but fixed (at
startup time) number of clients. Each RPC call from a client bears the client
number and a sequence number. \emph{Each execution core independently enforces
  execution of RPC calls in increasing sequence numbers}. This is done by
storing the last executed sequence number in a per-core durable datastructure as
part of the crash consistent transaction wrapping the execution of the RPC call
itself. Even in the presence of fail-overs, this durable memory of sequence
numbers seen from a client ensures at most once semantics. In addition to
guarding agains repeated execution of the same call, Cyclone also remembers the
\emph{results} of the last executed RPC call. This is done to assist clients
(particularly stateless clients). Cyclone allows the client to query the last
seen sequence number and result for that RPC call from the quorum of servers.

@@

We note that this design is not intended for performance, under the assumption
that distributed transactions will be infrequent compared to single key
operations.  On the other hand, it is worth comparing the lack of any need for
explicit fault recovery in Cyclone with fault recovery in systems systems that
explicity require such cleanup (~\cite{farm}, pp 59--63).  We believe that
such designs, which make availability easy for NVM applications are key to
increasing the adoption rate of NVM by data center operators.

@@

\section{Evaluation}
We evaluate Cyclone against its design goals by using it for up to 3-way
replication in a cluster of twelve 48 core Xeon servers. Three servers run
Cyclone's server side code, while the remaining machines run multithreaded
clients to load the system for our studies. The server side machines are each
equipped with four commodity 82599-based 10 Gbit ethernet ports, while each
client machine has one 82599-based ethernet port. All the machines are connected
through the same commodity 10 Gbit switch. We use 8 cores on the server machines
for running 8 instance of RAFT, leaving 39 cores to run application code. The
last remaining core is reserved for management code. We use NVML with the non
volatile pool held in DRAM, which proxies for the performance we would expect
from NVDIMMs. We evaluate Cyclone in three stages.

First, we evaluate pure replication performance by replicating RPC
calls. Cyclone simply replicates the contents of the RPC call without executing
it and returns a response to the client when replication is complete.
Figure~\ref{fig:pure_rep} shows the results. We consider two cases, one where
the clients send an RPC call with no payload and one where they add a 512 byte
dummy payload. For the no payload case the bottleneck is the message handling
rate. We note that Cyclone can handle small message replication rates as high as
\emph{8 million messages a second}. In the case of RPC calls with 512 byte
payloads, the bottleneck is the available network bandwidth. To show this, we
also plot transmit throughput derived from the message by adding the rate at
which data is sent to replicas from all the 8 quorum leaders to the rate at
which responses are sent back to the clients.  It is interesting to note that
with 3 replicas Cyclone achieves \emph{more than 40 Gbps} of bandwidth. The
reason for this is that the quorum leaders can end up residing on different
machines. Although replication traffic is still limited by the receive bandwidth
the responses to the clients, originating from different machines have access to
independent 4x10GigE ports, allowing total transmit throughput to exceed 40
Gbps. Although not discussed in this paper, Cyclone's control plane is designed
to exploit this by trying to spread quorum leaders across machines.

@@

Next, we consider the effect of adding NVMLs crash consistent
transactions. Figure~\ref{fig:cc_rep} shows the effect of adding a crash
consistent transaction where the application still does nothing but Cyclone
durably remembers the last sequence number and RPC call result (the application
simply returns the received payload). Throughput is now bottlenecked by NVML
allocator performance at just under a million messages a second. The final
result is adding the hash table in the picture with the clients updating the
values of keys already present in the hash table - a stable workload for easy
comparison. As Figure~\ref{fig:app_rep} shows performance stays at the just
under 1M calls a second mark for pure updates. For comparison we also show a
pure read only case where the clients only query the value of a key - there is
no replication \emph{nor} any crash consistent transaction applied. It is
interesting to note that the pure read only case is bottlenecked at the same
point as the pure replication case of 8 million operations a second.

These results lead us to conclude that Cyclone provides adequate performance
headroom for transparent replication in NVML applications. In fact our current
focus in the light of these results is towards improving application performance
on top of NVML. NVML is a general purpose library written for a variety of
substrates including traditional SSD based storage and there is adequate room
for improvement of performance for applications running on directly attached
persistent memory with close to DRAM performance.

>>>>

\subsection*{Abstract}
The introduction of non-volatile directly attached memory in datacenters - in
the form of NVDIMMs and new memory technologies allows efficient data structures
to be built on non-volatile heaps that provide good performance coupled with
durability. However high availability remains a problem for these primarily
single machine solutions. While transactional interfaces built over RDMA capable
networking hardware offer a viable alternative, such hardware is often
unavailable in datacenters. Cyclone offers a solution to this problem by
providing transparent replication middleware that can be used with Intel's Non
Volatile Memory Library (NVML) to construct durable data-structures that map
keys to values.  Cyclone is designed around an optimized userspace networking
stack to reduce latency and can scale its consensus protocol across cores and
network ports to provide good throughput under load. We show that Cyclone can
replicate as many as 7 million updates a second using just 4 commodity 10 GigE
ports on a machine.

@@

\section{Single Node Programming Model}
Cyclone assumes a concurrent durable application on a non-volatile heap as the
starting point. A number of libraries and filesystems make this possible
today~\cite{nvml, dax, pmfs, mnemosyne, nvheaps, cdds}. Cyclone currently
supports applications built on top of Intel's Non Volatile Memory Library
(NVML~\cite{nvml}). We describe NVML's programming model in this section to
provide adequate background for the rest of this paper. NVML requires 
that code that makes changes to the non-volatile heap is enclosed in a crash
consistent transaction such that either all of the transaction executes or none
of it executes in the event of power failures.
%This is critical to maintaining consistency as the data structure, being
%durable, must survive across power failures.
Figure~\ref{fig:example} illustrates the use of NVML to build a persistent
linked list. Updates to the linked list are enclosed in a NVML transaction to
avoid corruption due to power failure. It is also worth noting that NVM
libraries such as NVML usually allow remapping non-volatile heaps to a different
virtual memory address range across restarts (e.g., using {\tt TOID} macros to
implement fat pointers in NVML).

@@

Decoupled replication provides good performance and we believe it would be
the common case for RPC calls replicated via Cyclone. However the program might 
on occasion require successful replication to be a precondition to exection.
A case where we find this necessary is
\textit{output commit}~\cite{output_commit} - where the RPC call has
external side effects (e.g., communication with other machines in the cluster)
and therefore crash consistency is insufficient to roll back the execution of
the RPC call in the event that replication fails. We therefore provide users the
ability to declare at the client side that an RPC call should execute
in a coupled manner at the server side. Cyclone would then ensure that the RPC
call is successfully replicated before it attempts execution on any of the
replicas.

@@

We exploit the fact that DPDK allows prepending data to a packet
without physically moving the data in a packet buffer. The received packet
contains only the RPC call from the client. We prepend a RAFT related header
(containing the previous log index, previous log term, leader's term and
leader's commit point~\cite{raft}) to the packet buffer itself. To this RAFT header 
we then prepend an IPv4 header. We assign pre-determined values to the
IP address fields in order to exploit 5-tuple steering at the receiving NIC for 
the placement of the incoming packet in the correct queue.
This composite packet buffer is then sent \emph{unchanged} to every replica.
To properly route the packet through switches, for each follower replica we
allocate a small buffer containing an ethernet header with appropriate addresses
and \emph{chain} the buffer in front of the prepared packet - thereby ensuring
that we do not need to change the contents of the packet itself.

@@

Cyclone provides at most once semantics. It supports a large but fixed (at
startup time) number of clients. Each RPC call from a client bears the client
number and a sequence number. \emph{Each execution core independently enforces
  that accepted RPC calls from the same client bear increasing sequence
  numbers}. This is done by storing the last executed sequence number for every
client in a per-core durable datastructure as part of the crash consistent
transaction wrapping the execution of the RPC call. Even in the presence of
fail-overs, this durable memory of sequence numbers seen from a client ensures
at most once semantics. In addition to guarding against repeated execution of
the same call, Cyclone also remembers the \emph{result} of the last executed RPC
call from every client. This is done to assist clients (particularly stateless
clients). Cyclone allows the client to query the last seen sequence number and
result for that RPC call from the quorum of servers.

@@

We note that this design is not intended for performance, under the assumption
that distributed transactions will be infrequent compared to single key
operations.  On the other hand, it is worth comparing the lack of any need for
explicit fault recovery in Cyclone with fault recovery in systems systems that
explicity require such cleanup (~\cite{farm}, pp 59--63).  We believe that
such designs, which make availability easy for NVM applications are key to
increasing the adoption rate of NVM by datacenter operators.

@@

\section{Evaluation}
We evaluate Cyclone against its design goals by using it for up to 3-way
replication in a cluster of twelve 48 core Xeon servers. Three servers run
Cyclone's server side code, while the remaining machines run multithreaded
clients. The server side machines are each
equipped with four commodity 82599-based 10 Gbit ethernet ports, while each
client machine has one 82599-based 10 Gbit ethernet port. All the machines are
connected through the same 10 Gbit switch. We use 8 cores on the
server machines for running 8 instance of RAFT, leaving 39 cores to run
application code. The last remaining core is reserved for management code. We
use NVML with the non volatile pool held in DRAM, which proxies for the
performance we would expect from NVDIMMs. We evaluate Cyclone in three stages.

First, we evaluate pure replication performance by replicating RPC
calls. Cyclone simply replicates the contents of the RPC call without executing
it and returns a response to the client when replication is complete.
Figure~\ref{fig:pure_rep} shows the results. We consider two cases, one where
the clients send an RPC call with no payload and one where they add a 512 byte
dummy payload. For the no payload case the bottleneck is the message handling
rate. We note that Cyclone can handle small message replication rates as high as
\emph{7 million messages a second}. In the case of RPC calls with 512 byte
payloads, the bottleneck is the available network bandwidth. To show this, we
also plot transmit throughput derived from the message by adding the rate at
which data is sent to replicas from all the 8 quorum leaders to the rate at
which responses are sent back to the clients.  It is interesting to note that
with 3 replicas Cyclone achieves \emph{more than 40 Gbps} of bandwidth. The
reason for this is that the quorum leaders can end up residing on different
machines. Although replication traffic is still limited by the receive bandwidth
at follower replicas, the responses to the clients, originating from different
machines have access to independent 4x10GigE ports, allowing total transmit
throughput to exceed 40 Gbps. Although not discussed in this paper, Cyclone's
control plane is designed to exploit this by trying to spread quorum leaders
across machines.

@@

Next, Figure~\ref{fig:cc_rep} shows the effect of adding a crash
consistent transaction where the application still does nothing but Cyclone
durably remembers the last sequence number and RPC call result (the application
simply returns the received payload). Throughput is now bottlenecked by NVML
allocator performance at just under a million messages a second. The final
result is adding the hash table in the picture with the clients updating the
values of keys already present in the hash table - a stable workload for easy
comparison. As Figure~\ref{fig:app_rep} shows performance stays at the just
under 1M calls a second mark for pure updates. We also show a
pure read only case where the clients only query the value of a key - there is
no replication \emph{nor} any crash consistent transaction applied. It is
interesting to note that the pure read only case is bottlenecked almost at the same
point as the pure replication case of 7 million operations a second.

These results lead us to conclude that Cyclone provides adequate performance
headroom for transparent replication in NVML applications. In fact our current
focus in the light of these results is towards improving application performance
on top of NVML rather than improving the performane of replication in Cyclone.
%NVML is a general purpose library written for a variety of
%substrates including traditional SSD based storage and there is adequate room
%for improvement of performance for applications running on directly attached
%persistent memory with close to DRAM performance.
